Centers and centralizers of higher enveloping algebras of Lie algebras

### 2020-10-02 draft.tex
def

\section{A few examples of the central result}

We apply our central result to important classes of Lie algebras: 
abelian, semisimple, and free.

Given a free dg commutative algebra $A = \Sym(V^\bullet)$ generated by a cochain complex $V^\bullet$,
one can view it as an $\EE_n$ algebra via the forgetful functor determined by the operad map $\EE_n \to E_\infty$.
We will blithely view any commutative algebra as an $\EE_n$ algebra for any $n$ via this forgetting.

Using our models, we obtain a generalization of the Harish--Chandra isomorphism to the $\EE_n$ center of the  enveloping $\EE_n$ algebra. 
The following result follows from a simple application of the Poincar\'{e} lemma for de Rham forms and compactly supported de Rham forms.

\begin{prp}
There is a canonical quasi-isomorphism of $\EE_{n+1}$-algebras
\[
\bZ_n(\UU_n \fg) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\]
\end{prp}

\begin{prp}
For $\fg$ a semi-simple Lie algebra with Cartan subalgebra~$\fh$ and Weyl group~$W$, we have the following:
\begin{itemize}
\item if $n$ is odd, then there is a quasi-isomorphism of $\EE_{n+1}$-algebras
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies (\fg) \otimes \Sym(\fh[1-n])^W.
\]
\item if $n$ is even, then there is a quasi-isomorphism of $\EE_{n+1}$-algebras
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies (\fg) \otimes \Tilde{H}_\bu^{\rm Lie} \left(\fg, \frac{n}{2} -1\right).
\]
\end{itemize}
\end{prp}
\begin{proof}
By the Poincar\'{e} lemma for de Rham forms and compactly supported de Rham forms, one has a canonical quasi-isomorphism
\[
\bZ_n(\UU_n \fg) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\]
Since $\fg$ is semi-simple, there is a quasi-isomorphism
\[
\clies(\fg, \Sym(\fg[1-n]) \simeq \clies (\fg) \otimes \Sym(\fg[1-n])^G 
\]
where $G$ is a compact connected Lie group whose Lie algebra is $\fg$.

@@

The universal property of centralizer (recall definition~\ref{D:centralizer}) implies that 
for any maps $A\stackrel{f}\to B$ and $B\stackrel{g}\to C$ of $\EE_n$ algebras, 
there is a natural map of $\EE_n$ algebras
\begin{equation}
\label{eqdef:canonicaltocentralizers} 
{\frak Z}_n(g) \otimes {\frak Z}_n(f) \longrightarrow {\frak Z}_n(g\circ f)
\end{equation}
that arises from following commutative diagram of $\EE_n$ algebras:
\[
\begin{tikzcd}
&   {\frak Z}_n(g) \otimes  {\frak Z}_n(f) \otimes A \arrow[rr, "\id\otimes \ev_f"] &&  {\frak Z}_n(g) \otimes  B \arrow[dr, "\ev_g"]\\
A \arrow[ur, "1\otimes \id_A"] \arrow[rr, "f"] && B \arrow[rr, "g"] \arrow[ur, "1\otimes \id_B"] && C 
\end{tikzcd}.
\]
When $f=g=\id_A:A\to A$, we thus obtain a canonical morphism 
\begin{equation} 
\label{eq:dgAssofcenter}
{\frak Z}_n(A){\otimes} {\frak Z}_n(A) \longrightarrow {\frak Z}_n(A).
\end{equation}
Lurie \cite{HA} has proved that this multiplication makes the center into an $E_1$ algebra in $\EE_n$ algebras 
and therefore an $E_{n+1}$ algebra, by Dunn additivity. 
%To be precise, there is the following result.
%
%\begin{prp}[\cite{LurieHA} Section 5] 
%\label{P:PtyCenter}
%\begin{itemize}
% \item The canonical map~\eqref{eq:dgAssofcenter} makes the center ${\frak Z}_n(A)$ into an $E_1$-algebra in the symmetric monoidal category of 
% $\EE_n$-algebras.
% \item There is a canonical equivalence of the $\infty$-category $E_1\mathrm{-Alg}(\EE_n\mathrm{-Alg})$ with $E_{n+1}\mathrm{-Alg}$.
%\end{itemize}
%\end{prp}

@@

>>>>

@@

The universal property of centralizer (recall definition~\ref{D:centralizer}) implies that 
for any maps $A\stackrel{f}\to B$ and $B\stackrel{g}\to C$ of $\EE_n$ algebras, 
there is a natural map of $\EE_n$ algebras
\begin{equation}
\label{eqdef:canonicaltocentralizers} 
{\frak Z}_n(g) \otimes {\frak Z}_n(f) \longrightarrow {\frak Z}_n(g\circ f)
\end{equation}
that arises from following commutative diagram of $\EE_n$ algebras:
\[
\begin{tikzcd}
&   {\frak Z}_n(g) \otimes  {\frak Z}_n(f) \otimes A \arrow[rr, "\id\otimes \ev_f"] &&  {\frak Z}_n(g) \otimes  B \arrow[dr, "\ev_g"]\\
A \arrow[ur, "1\otimes \id_A"] \arrow[rr, "f"] && B \arrow[rr, "g"] \arrow[ur, "1\otimes \id_B"] && C 
\end{tikzcd}.
\]
When $f=g=\id_A:A\to A$, we thus obtain a canonical morphism 
\begin{equation} 
\label{eq:dgAssofcenter}
{\frak Z}_n(A){\otimes} {\frak Z}_n(A) \longrightarrow {\frak Z}_n(A).
\end{equation}
Lurie \cite{HA} has proved that this multiplication makes the center into an $\EE_1$ algebra in $\EE_n$ algebras 
and therefore an $\EE_{n+1}$ algebra, by Dunn additivity. 
%To be precise, there is the following result.
%
%\begin{prp}[\cite{LurieHA} Section 5] 
%\label{P:PtyCenter}
%\begin{itemize}
% \item The canonical map~\eqref{eq:dgAssofcenter} makes the center ${\frak Z}_n(A)$ into an $E_1$-algebra in the symmetric monoidal category of 
% $\EE_n$-algebras.
% \item There is a canonical equivalence of the $\infty$-category $E_1\mathrm{-Alg}(\EE_n\mathrm{-Alg})$ with $E_{n+1}\mathrm{-Alg}$.
%\end{itemize}
%\end{prp}

@@

\section{A few applications and examples of the central result}

In this section we pursue a few immediate corollaries of the central results above.
We apply our central result to important classes of Lie algebras: 
abelian, semisimple, and free.

\subsection{Deformations of $\EE_n$ enveloping algebras}
It is well-known that the Hochschild cohomology of an associative, or $\EE_1$-algebra, controls its deformations. 
Precisely, if $A$ is an $\EE_1$-algebra then the shift of its Hochschild cohomology ${\rm Hoch}^\bu (A , A)[1]$ carries a Lie algebra structure (in the homotopical sense). 
The Maurer--Cartan elements of this dg Lie algebra classify deformations of $A$ as an $\EE_1$-algebra.

Similarly, if $A$ is an $\EE_n$-algebra, the $\EE_n$-center $\fZ_n(A)$ controls deformations of $A$ as $\EE_n$-algebra. 
Since $\fZ_n(A)$ is an $\EE_{n+1}$-algebra, the shift $\fZ_n(A)[n]$ is equipped with a Lie bracket. 
Maurer--Cartan elements classify deformations of $A$ as an $\EE_{n}$-algebra.

Using our model for the $\EE_n$-center, we obtain an explicit description 
of deformations of $\EE_n$-enveloping algebras.

\begin{prp}\label{prop:def}
There is a canonical quasi-isomorphism 
\begin{equation}\label{eqn:def}
\bZ_n(\UU_n \fg) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\end{equation}
%In particular, the cochain complex on the right-hand side has the structure of an $\EE_{n+1}$ algebra. 
In particular, $\EE_n$-deformations of $\UU_n \fg$ up to equivalence are in bijective correspondence with the cohomology group
\[
H^{n+1} \bigg(\clie^\bu \left(\fg , \Sym(\fg[1-n]) \right) \bigg).
\]
\end{prp}
\begin{proof}
The first statement follows from the Poincar\'{e} lemma for de Rham forms and compactly supported de Rham forms
\[
\bZ_n(\UU_n \fg) = \clie^\bu\left(\Omega^\bu(\RR^n) \otimes \fg, \clieu_\bu(\Omega^\bu_c (\RR^n) \otimes \fg\right) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\]
The second statement follows from the fact that Maurer--Cartan elements in $\bZ_n(\UU_n \fg)[n]$ classify deformations of $\UU_n \fg$. 
\end{proof}

\begin{rmk}
When $n=1$ the quasi-isomorphism of the above proposition produces the well-known quasi-isomorphism
\[
{\rm Hoch}^\bu(U \fg , U \fg) \simeq \clie^\bu(\fg , \Sym(\fg)) ,
\]
see \cite{LodayCyclicBook}. 
\end{rmk}

%As a corollary of the above proposition, we obtain a concrete description for $\EE_n$-deformations of $\UU_n \fg$ when $\fg$ is an ordinary (strict) Lie algebra. 
Suppose $\fg$ is an ordinary Lie algebra.
It is known that if $\fg$ is semi-simple then the $\EE_1$-enveloping algebra $\UU_1 \fg$ is rigid in the sense that it admits no nontrivial deformations. 
This follows from the above equivalence (\ref{eqn:def}) together with the fact that $H_{\rm Lie}^2(\fg , \Sym(\fg)) = 0$ when $\fg$ is semi-simple by Whitehead's theorem.

For higher $n$, the $\EE_n$-enveloping algebra is no longer rigid, even in the case of semi-simple Lie algebras. 
For instance, consider the case $n=2$.
\begin{cor}
Deformations of $\UU_2 \fg$ are in bijective correspondence with the group
\[
H^3_{\rm Lie}(\fg) \oplus H^2_{\rm Lie}(\fg , \fg) \oplus H^1_{\rm Lie}(\fg , \wedge^2 \fg) \oplus \wedge^3(\fg)^\fg .
\]
\end{cor}

In particular, when $\fg$ is semisimple the middle two terms in the above corollary vanish and we are left with a two-dimensional space of deformations $H^3_{\rm Lie}(\fg) \oplus \wedge^3 (\fg)^\fg = k \oplus k$. \brian{check me}

We have not yet described explicitly the Lie algebra structure on $\bZ_n(\UU_n \fg)[n]$ which is implicitly used in the characterization of deformations above. 
\brian{can we describe $\PP_{n+1}$ structure explicitly?}

Given a free dg commutative algebra $A = \Sym(V^\bullet)$ generated by a cochain complex $V^\bullet$,
one can view it as an $\EE_n$ algebra via the forgetful functor determined by the operad map $\EE_n \to \EE_\infty$.
We will blithely view any commutative algebra as an $\EE_n$ algebra for any $n$ via this forgetting.

Using our models, we can obtain a generalization of the Harish--Chandra isomorphism to the $\EE_n$ center of the  enveloping $\EE_n$ algebra. 
%The following result follows from a simple application of the Poincar\'{e} lemma for de Rham forms and compactly supported de Rham forms.

\begin{prp}
For $\fg$ a semi-simple Lie algebra with Cartan subalgebra~$\fh$ and Weyl group~$W$, we have the following:
\begin{itemize}
\item if $n$ is odd, then there is a quasi-isomorphism 
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies (\fg) \otimes \Sym(\fh[1-n])^W.
\]
\item if $n$ is even, then there is a quasi-isomorphism 
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies (\fg) \otimes \Tilde{H}_\bu^{\rm Lie} \left(\fg, \frac{n}{2} -1\right).
\]
\end{itemize}
\end{prp}
\begin{proof}
We apply the quasi-isomorphism of Proposition \ref{prop:def}
\[
\bZ_n(\UU_n \fg) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\]
Since $\fg$ is semi-simple, there is a further quasi-isomorphism
\[
\clies(\fg, \Sym(\fg[1-n]) \simeq \clies (\fg) \otimes \Sym(\fg[1-n])^G 
\]
where $G$ is a compact connected Lie group whose Lie algebra is $\fg$.
